---
title: "Homework 4: Diffusion of Tetracycline"
output: pdf_document
---

We continue examining the diffusion of tetracycline among doctors in Illinois in the early 1950s, building on our work in lab 6.  You will need the data sets `ckm_nodes.csv` and `ckm_network.dat` from the labs.

1. Clean the data to eliminate doctors for whom we have no adoption-date information, as in the labs.  Only use this cleaned data in the rest of the assignment.
```{r,warning=FALSE}
library(tidyverse)
setwd("D:/github/Rlab/")
ckm_nodes<-read.csv("data/ckm_nodes.csv",header=T)
ckm_network<-read.table("data/ckm_network.dat")

ckm_network = ckm_network[which(!is.na(ckm_nodes$adoption_date))
                          ,which(!is.na(ckm_nodes$adoption_date))]
ckm_nodes <- ckm_nodes %>%
    filter(!is.na(adoption_date))
```


2. Create a new data frame which records, for every doctor, for every month, whether that doctor began prescribing tetracycline that month, whether they had adopted tetracycline before that month, the number of their contacts who began prescribing strictly _before_ that month, and the number of their contacts who began prescribing in that month or earlier.  Explain why the dataframe should have 6 columns, and 2125 rows.  Try not to use any loops.
```{r}
new = data.frame(matrix(rep(c(1:125),17),nrow=125))
new = stack(new)
names(new)<-c("doc","adoption_date")
new$adoption_date = substr(new$adoption_date,2,3)
new$adoption_date[which(new$adoption_date==17)] = Inf

adoption_date_net = data.frame(t(matrix(rep(ckm_nodes$adoption_date,125),nrow=125)))
compare_small = adoption_date_net<t(adoption_date_net)
compare_notbig = adoption_date_net<=t(adoption_date_net)

compare_small_net = compare_small&ckm_network
compare_notbig_net = compare_notbig&ckm_network

small = rowSums(as.matrix(compare_small_net))
notbig = rowSums(as.matrix(compare_notbig_net))

new <- new %>% mutate(if_start= ifelse(ckm_nodes$adoption_date[doc]==adoption_date,TRUE,FALSE))%>%
    mutate(if_before = ifelse(ckm_nodes$adoption_date[doc]<adoption_date,TRUE,FALSE))%>%
    mutate(connect_before = rep(small,17))%>%
    mutate(connect_not_after = rep(notbig,17))
    
```

3.  Let
    \[
    p_k = \Pr(\text{A doctor starts prescribing tetracycline this month} \mid \\
    \text{Number of doctor's contacts prescribing before this month}=k)
    \]
    and
   \[
     q_k = \Pr(\text{A doctor starts prescribing tetracycline this month} \mid \\ \text{Number of doctor's contacts prescribing this month}=k)
    \]
   We suppose that $p_k$ and $q_k$ are the same for all months.
     a. Explain why there should be no more than $21$ values of $k$ for which we can estimate $p_k$ and $q_k$ directly from the data.
```{r}
max(rowSums(ckm_network))
```

explain: because a doctor has at most 20 connections, thus k must no more than 20, there are 21 choices.

b. Create a vector of estimated $p_k$ probabilities, using the data frame from (2).  Plot the probabilities against the number of prior-adoptee contacts $k$.

```{r,warning=FALSE}
doctor_month_before = matrix(nrow = 125,ncol = 17)
for(i in 1:125){
    for(j in 1:17){
        doctor_month_before[i,j] =
          length(which(ckm_nodes$adoption_date[which(ckm_network[i,]==1)]<j))
    }
}
p_pro = rep(0,max(doctor_month_before)+1)
doc_per_situation = rep(0,max(doctor_month_before)+1)
for(i in 0:max(doctor_month_before)){
    all = length(which(doctor_month_before==i))#all the cases 
    #when a doctor have i conections have take pills
    num_start = 0# all the cases of a doctor starts take pills 
    #when i connections started to take pills before.
    for(j in 1:125){
        if(ckm_nodes$adoption_date[j]!=Inf &
           doctor_month_before[j,ckm_nodes$adoption_date[j]]==i){
            num_start = num_start + 1
            doc_per_situation[i] = doc_per_situation[i] + 1
        } 
    }
    p_pro[i] = num_start/all
}
p_pro = na.omit(p_pro)

ggplot(data = data.frame(p_pro)) +  
  geom_point(aes(x = 1:length(p_pro), y = p_pro)) + 
  geom_line(aes(x = 1:length(p_pro), y = p_pro),colour='red') +
  labs(title = "p_k~k",   y = "p_k",   x = "k") +
  theme_bw()
```

     c. Create a vector of estimated $q_k$ probabilities, using the data frame from (2). Plot the probabilities against the number of prior-or-contemporary-adoptee contacts $k$.
```{r,warning=FALSE}
doctor_month_not_after = matrix(nrow = 125,ncol = 17)
for(i in 1:125){
    for(j in 1:17){
        doctor_month_not_after[i,j] =
          length(which(ckm_nodes$adoption_date[which(ckm_network[i,]==1)]<=j))
    }
}

q_pro = rep(0,max(doctor_month_not_after)+1)
for(i in 0:max(doctor_month_not_after)){
    all = length(which(doctor_month_not_after==i))#all the cases
    #when a doctor have i conections have take pills
    num_start = 0# all the cases of a doctor start to take pills
    #when i connections started to take pills before.
    for(j in 1:125){
        if(ckm_nodes$adoption_date[j]!=Inf &
           doctor_month_not_after[j,ckm_nodes$adoption_date[j]]==i){
            num_start = num_start + 1
        } 
    }
    q_pro[i+1] = num_start/all
}
q_pro = na.omit(q_pro)
ggplot(data = data.frame(q_pro)) +  
  geom_point(aes(x = 0:(length(q_pro)-1), y = q_pro)) + 
  geom_line(aes(x = 0:(length(q_pro)-1), y = q_pro),colour='red') +
  labs(title = "q_k~k",   y = "q_k",   x = "k") +
  theme_bw()
```
     

4. Because it only conditions on information from the previous month, $p_k$ is a little easier to interpret than $q_k$.  It is the probability per month that a doctor adopts tetracycline, if they have exactly $k$ contacts who had already adopted tetracycline.
    a. Suppose $p_k = a + bk$.  This would mean that each friend who adopts the new drug increases the probability of adoption by an equal amount.  Estimate this model by least squares, using the values you constructed in (3b).  Report the parameter estimates.
```{r}
#since when k>=10 and k=7, the pro is all zero or NA because of the lack of data,
#we delete these probabilities.
x = c(1:6,8,9)
lm.fit = lm(p_pro[x]~x)
summary(lm.fit)
```
    
    b. Suppose $p_k = e^{a+bk}/(1+e^{a+bk})$.  Explain, in words, what this model would imply about the impact of adding one more adoptee friend on a given doctor's probability of adoption.  (You can suppose that $b > 0$, if that makes it easier.)  Estimate the model by least squares, using the values you constructed in (3b).
```{r}
#adding one more adoptee friend will make a given doctor's probability of 
#adoption bigger and closer to 1, but not bigger than 1(when b>0)
mse1<-function(param,x_=x,y=p_pro[x]){
    a = param[1]
    b = param[2]
    estimate = exp(a+b*x_)/(1+exp(a+b*x_))
    MSE = mean((y-estimate)^2)
    return(MSE)
}
res = nlm(mse1, c(a=0.1,b=0.1))
res
```
    estimate = exp(a+b*x)/(1+exp(a+b*x))
    MSE = mean((y-estimate)^2)
    return(MSE)
    
c. Plot the values from (3b) along with the estimated curves from (4a) and (4b).  (You should have one plot, with $k$ on the horizontal axis, and probabilities on the vertical axis .)  Which model do you prefer, and why?

```{r}
lm.predict1 = predict(lm.fit)
p.estimate = exp(res$estimate[1]+res$estimate[2]*x)/(1+exp(res$estimate[1]+res$estimate[2]*x))
ggplot(data = data.frame(p_pro[x])) +  
  geom_point(aes(x = x, y = p_pro[x])) + 
  geom_smooth(aes(x = x, y = lm.predict1),colour='red') +
  geom_smooth(aes(x = x, y = p.estimate)) +
  labs(title = "pro~k",   y = "pro",   x = "k") +
  theme_bw()
```
I prefer the second model, because the probability can not exist 1.

_For quibblers, pedants, and idle hands itching for work to do_: The $p_k$ values from problem 3 aren't all equally precise, because they come from different numbers of observations.  Also, if each doctor with $k$ adoptee contacts is independently deciding whether or not to adopt with probability $p_k$, then the variance in the number of adoptees will depend on $p_k$.  Say that the actual proportion who decide to adopt is $\hat{p}_k$.  A little probability (exercise!) shows that in this situation, $\mathbb{E}[\hat{p}_k] = p_k$, but that $\mathrm{Var}[\hat{p}_k] = p_k(1-p_k)/n_k$, where $n_k$ is the number of doctors in that situation.  (We estimate probabilities more precisely when they're really extreme [close to 0 or 1], and/or we have lots of observations.)  We can estimate that variance as $\hat{V}_k = \hat{p}_k(1-\hat{p}_k)/n_k$.  Find the $\hat{V}_k$, and then re-do the estimation in (4a) and (4b) where the squared error for $p_k$ is divided by $\hat{V}_k$.  How much do the parameter estimates change?  How much do the plotted curves in (4c) change?
```{r}
Vk = p_pro[x]*(1-p_pro[x])/doc_per_situation[x]
#4(a)
mse1<-function(param,x_=x,y=p_pro[x],V =Vk){
    a = param[1]
    b = param[2]
    estimate = a+x_*b
    MSE = mean((y-estimate)^2/V)
    return(MSE)
}
res1 = nlm(mse1, c(a=0.1,b=0.1))
#4(b)
mse2<-function(param,x_=x,y=p_pro[x],V =Vk){
    a = param[1]
    b = param[2]
    estimate = exp(a+b*x_)/(1+exp(a+b*x_))
    MSE = mean((y-estimate)^2/V)
    return(MSE)
}
res2 = nlm(mse2, c(a=-3,b=0.04))
#4(c)
p.estimate1 = res1$estimate[1]+res1$estimate[2]*x
p.estimate2 = exp(res2$estimate[1]+res2$estimate[2]*x)/(1+exp(res2$estimate[1]+res2$estimate[2]*x))

ggplot(data = data.frame(p_pro[x])) +  
  geom_point(aes(x = x, y = p_pro[x])) + 
  geom_smooth(aes(x = x, y = p.estimate1),colour='red') +
  geom_smooth(aes(x = x, y = p.estimate2)) +
  labs(title = "pro~k",   y = "pro",   x = "k") +
  theme_bw()
```
The parameters of linier model change from 0.0377,0.00137 to 0.0587,-0.00518; the parameters of exponential model change from -3,30,0.0452 to -2.62,-0.178.

The fitting line in the plot are more aline with the first 4 dots.
