---
title: 'Lab 3: Data Wrangling on Soccer Tournament Data'
date: 'July 9, 2020'
output:
  pdf_document: default
  html_document: default
---
```{r setup, message = F, include=FALSE}
library(tidyverse)
```

Soccer tournament data wrangling
--------------------------
Read the dataset of football games.
```{r,warning=FALSE}
setwd("D:/github/Rlab/")
d <- read_csv("data/results.csv")
```

1. Select variables `date`, `home_team` and `away_team`.
```{r}
d1 = d %>% dplyr::select(date, home_team, away_team)
```
2. Subset games with **Brazil** as the home team.
```{r}
d2 = d1 %>% dplyr::filter(home_team == "Brazil")
```

3. Choose the games that Brazil won as the home team, and select variables `date`, `away_team` and `tournament`.
```{r}
d3 = d %>% dplyr::filter(home_team == "Brazil") %>%
  dplyr::filter(home_score>away_score) %>%
  dplyr::select(date, home_team, tournament)
```

4. Add the difference of goals, and an indicator variable called `goleada` for when the difference of goals is large,  and select what we did only for Brazil. **Hint: use `ifelse`**.
```{r}
d4 = d %>% dplyr::mutate(diff_goal = home_score-away_score)%>%
  dplyr::mutate(goleada = ifelse(diff_goal >= 5|diff_goal<=-5, 1, 0)) %>%
  dplyr::filter(home_team == "Brazil"|away_team == "Brazil")
```

5. What was the largest difference in goals within these games?
```{r}

```

6. The top 5 goleadas?
```{r}

```
7. Summary on goals scored by home teams, such as `mean` of `home_score` and `away_score`, `std`, using `group_by` and `summarise`
```{r}
d %>% dplyr::group_by(home_team)%>%
  summarize(mean_homescore = mean(home_score))%>%
  ungroup()
d %>% dplyr::group_by(home_team)%>%
  summarize(cont_awayscore = mean(away_score))%>%
  ungroup()
d %>% dplyr::group_by(home_team)%>%
  summarize(std_homescore = sd(home_score))%>%
  ungroup()
```

8. Proportion of victories of **Brazil** on different tournaments against each opponent, for instance, **Argentina**.
```{r}
d1 = d %>% dplyr::filter(home_team == "Brazil"|away_team == "Brazil")%>%
  dplyr::mutate(against = ifelse(home_team == "Brazil",away_team,home_team))%>%
  dplyr::mutate(all = 1)%>%
  dplyr::mutate(win = ifelse(home_team == "Brazil",ifelse(home_score>away_score,1,0),ifelse(home_score<away_score,1,0)))


d1%>% group_by(tournament,against)%>%
  summarize(win_rate = sum(win)/sum(all))
```
